---
title: "Code Topic 03 Team 03"
author: "Bolz, C.,Bonsen, M.,Pott, M., Simon, M."
date: "`r Sys.Date()`"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Clean-Up

## Load Libraries
```{r Load libraries}
library(tidyverse)
library(dplyr)
library(corrplot)
library(cluster) 
library(factoextra)
library(ggpubr)
library(readxl)
library(FactoMineR)
library(ggplot2)
```

## Load Raw Data

First, the experimental data is added.
```{r  Load Experiment Data}
Raw_data <- read.csv("https://raw.githubusercontent.com/datascience-mobi-2022/2022-topic-03-team-03/3a0f2d9583b56a1cd22150b399d61352d6f7b0e1/RDeeP_HeLa_Interphase.csv", header=TRUE, row.names=1, sep = ";")
```

Furthermore, we added data on previously found RBPs and non-RBPs.
```{r Load extra data for comparison of results}
HS_RBPs <- read.csv("https://raw.githubusercontent.com/datascience-mobi-2022/2022-topic-03-team-03/main/1_HS_RBPs.csv", header=TRUE, sep = ",")

HS_non_RBPs <- read.csv("https://raw.githubusercontent.com/datascience-mobi-2022/2022-topic-03-team-03/main/2_HS_non_RBPs.csv", header=TRUE, sep = ",")
```

## Reorganize Data

We split the data into the two samples: RNase treated and control group.
```{r Split Ctrl and RNase}
# Control
HeLa_Ctrl = Raw_data %>% select(contains("Ctrl")) 

# RNase
HeLa_RNase = Raw_data %>% select(contains("RNA"))
```

## Remove Rows of Zero

For some proteins no data was found (rows of zeroes). These proteins are excluded from the analysis.
```{r Rows of zero}
# Find rows of zeroes
trash <- c(which(rowMeans(HeLa_Ctrl) == 0),which(rowMeans(HeLa_RNase) == 0))
trash <- trash[-5]

# Remove failed proteins
HeLa_cleaned <- Raw_data[- trash, ]

# Separate data sets

# Control
HeLa_Ctrl = HeLa_cleaned %>% select(contains("Ctrl")) 
# RNase
HeLa_RNase = HeLa_cleaned %>% select(contains("RNA"))
```

# Normalization

## Normalize columns

For every fraction and every protein, we adjusted the total value of all three reps up to the same value. This step makes the individually obtained repeats of the experiment more comparable. 
```{r Normalize columns functions}
# 1. Create function that calculates the sum of the colums for each fraction for all three replicates
Spalten_Norm <- function(x){
  ColSumme = colSums(x)
  maxcolsums = c()
  k = 3
  for (i in 1:(ncol(x)/3)) {
    a = ColSumme[[k*i]]
    b = ColSumme[[k*i-1]]
    c = ColSumme[[k*i-2]]
    maxcolsums = c(maxcolsums, max(a,b,c))
  }
  return(ColSumme)
}
# 2. Create function that selects the maxcolsum for each fraction out of all three replicates
Spalten_Norm_1 <- function(x){
  ColSumme = colSums(x)
  maxcolsums = c()
  k = 3
  for (i in 1:(ncol(x)/3)) {
    a = ColSumme[[k*i]]
    b = ColSumme[[k*i-1]]
    c = ColSumme[[k*i-2]]
    maxcolsums = c(maxcolsums, max(a,b,c))
  }
  return(maxcolsums)
}
# 3. Create function that calculates the quotient of maxcolsum and colsum
Spalten_Norm_2 <- function(x, ColSumme, maxcolsums){
  Quotient = c()
  k = 3
  for (i in 1:(ncol(x)/3)){
    Quotient = c(Quotient, maxcolsums[i]/ColSumme[[k*i-2]])
    Quotient = c(Quotient, maxcolsums[i]/ColSumme[[k*i-1]])
    Quotient = c(Quotient, maxcolsums[i]/ColSumme[[k*i]])
  }
  return(Quotient)
}
```

Then, we applied our set of functions.
```{r Normalize columns}
# Apply created functions on our data set to calculate the normalized protein values per fraction and replicate
ColSumme<- Spalten_Norm(HeLa_cleaned)
maxcolsums <- Spalten_Norm_1(HeLa_cleaned)
Quotient <- Spalten_Norm_2(HeLa_cleaned, ColSumme, maxcolsums)
Hela_Col_Norm <- sweep(HeLa_cleaned, 2, Quotient, "*")

# Remove excess objects
rm(ColSumme, maxcolsums, Quotient)
```

## Normalize rows

Next, we brought the arbitrary values to a common percentage scale. For that purpose, we normalized the rows (by protein). The resulting values represent how much content of the total protein content was present in each fraction.
```{r Normalize rows}
# Sort data set to select individual replicates 
# Control
HeLa_Ctrl_colnorm = Hela_Col_Norm %>% select(contains("Ctrl")) 
HeLa_Ctrl1 = HeLa_Ctrl_colnorm %>% select(contains("Rep1"))
HeLa_Ctrl2 = HeLa_Ctrl_colnorm %>% select(contains("Rep2"))
HeLa_Ctrl3 = HeLa_Ctrl_colnorm %>% select(contains("Rep3"))
# RNase
HeLa_RNase_colnorm = Hela_Col_Norm %>% select(contains("RNA"))
HeLa_RNase1 = HeLa_RNase_colnorm %>% select(contains("Rep1"))
HeLa_RNase2 = HeLa_RNase_colnorm %>% select(contains("Rep2"))
HeLa_RNase3 = HeLa_RNase_colnorm %>% select(contains("Rep3"))

# Create function that scales the protein amount in each replicate per fraction to 100
 rowPercent <- function(dataframe){
     dt <- dataframe
     lastrow <- dim(dataframe)[1]
     for(i in 1:lastrow){
         RowSum <- rowSums (dataframe[i,])
    if (RowSum == 0) {dt[i,] <- 0}
    else dt[i,] <- (dataframe[i,]/RowSum)*100    } 
     return(dt)}
 
# Apply created function
Norm_HeLa_Ctrl1 <- rowPercent(HeLa_Ctrl1)
Norm_HeLa_Ctrl2 <- rowPercent(HeLa_Ctrl2)
Norm_HeLa_Ctrl3 <- rowPercent(HeLa_Ctrl3)

Norm_HeLa_RNase1 <- rowPercent(HeLa_RNase1)
Norm_HeLa_RNase2 <- rowPercent(HeLa_RNase2)
Norm_HeLa_RNase3 <- rowPercent(HeLa_RNase3)

Norm_HeLa <- cbind(Norm_HeLa_Ctrl1, Norm_HeLa_Ctrl2, Norm_HeLa_Ctrl3, Norm_HeLa_RNase1, Norm_HeLa_RNase2, Norm_HeLa_RNase3)
```

## Create separate dataframes

Next, data frames for each individual fraction and treatment option are created, containing the triplicate values for each protein. 

They follow the naming convention: 
- Fraction1_Ctrl for the triplicate values of the first fraction of the control group
- Fraction_8_RNase for the triplicate values of the eighth fraction of the RNase treated sample

```{r  Fractions}
#select the fractions
vector <- c(1, 26, 51, 76, 101, 126, 2, 27, 52, 77, 102, 127, 3, 28, 53, 78, 103, 128, 4, 29, 54, 79, 104, 129, 5, 30, 55, 80, 105, 130, 6, 31, 56, 81, 106, 131, 7, 32, 57, 82, 107, 132, 8, 33, 58, 83, 108, 133, 9, 34, 59, 84, 109, 134, 10, 35, 60, 85, 110, 135, 11, 36, 61, 86, 111, 136, 12, 37, 62, 87, 112, 137, 13, 38, 63, 88, 113, 138, 14, 39, 64, 89, 114, 139, 15, 40, 65, 90, 115, 140, 16, 41, 66, 91, 116, 141, 17, 42, 67, 92, 117, 142, 18, 43, 68, 93, 118, 143, 19, 44, 69, 94, 119, 144, 20, 45, 70, 95, 120, 145, 21, 46, 71, 96, 121, 146, 22, 47, 72, 97, 122, 147, 23, 48, 73, 98, 123, 148, 24, 49, 74, 99, 124, 149, 25, 50, 75, 100, 125, 150) 
HeLa_Fractions  <- Norm_HeLa[vector]

rm(vector)

# Create list of all separate data frames for the fractions
fractions_names <-  sapply(c(1:50), function(x) {
  if (x%%2 == 0) 
    str_glue("Fraction", (x/2), "_RNase")
  else 
    str_glue("Fraction", ((x+1)/2), "_Ctrl")
})

# Assign the names to the according fractions
for (i in c(1:50)) {
  assign(fractions_names[i], HeLa_Fractions[,(3*i-2):(3*i)])
}
```

## Calculate means of triplicates with one sigma cut-off

When calculating the mean value of a triplicate, sometimes outlier values in a triplicate obscured the likely true mean value for a fraction. Therefore, we excluded all values in a triplicate that differed more then one sigma from the calculated mean of all three replicates.  

```{r Mean}
# Create function to calculate mean and exclude triplicate values failing one sigma cut-off
mean_sd <- function(x) {
        m <- mean(x)
        s <- sd(x)
       for (i in 1:3)
        if (( (m - 1 * s) > x[i]) || ((m + 1 * s) < x[i])) {x[i] <- NA} # NA for outliers
 mean(x, na.rm =TRUE) }

# Apply created function to data
mean_Ctrl1 <- apply(Fraction1_Ctrl, 1,mean_sd)
mean_Ctrl2 <- apply(Fraction2_Ctrl, 1,mean_sd )
mean_Ctrl3 <- apply(Fraction3_Ctrl, 1,mean_sd )
mean_Ctrl4 <- apply(Fraction4_Ctrl, 1,mean_sd )
mean_Ctrl5 <- apply(Fraction5_Ctrl, 1,mean_sd)
mean_Ctrl6 <- apply(Fraction6_Ctrl, 1,mean_sd)
mean_Ctrl7 <- apply(Fraction7_Ctrl, 1,mean_sd)
mean_Ctrl8 <- apply(Fraction8_Ctrl, 1,mean_sd)
mean_Ctrl9 <- apply(Fraction9_Ctrl, 1,mean_sd)
mean_Ctrl10 <- apply(Fraction10_Ctrl, 1, mean_sd)
mean_Ctrl11 <- apply(Fraction11_Ctrl, 1,mean_sd)
mean_Ctrl12 <- apply(Fraction12_Ctrl, 1,mean_sd)
mean_Ctrl13 <- apply(Fraction13_Ctrl, 1,mean_sd)
mean_Ctrl14 <- apply(Fraction14_Ctrl, 1,mean_sd)
mean_Ctrl15 <- apply(Fraction15_Ctrl, 1,mean_sd)
mean_Ctrl16 <- apply(Fraction16_Ctrl, 1,mean_sd)
mean_Ctrl17 <- apply(Fraction17_Ctrl, 1,mean_sd)
mean_Ctrl18 <- apply(Fraction18_Ctrl, 1,mean_sd)
mean_Ctrl19 <- apply(Fraction19_Ctrl, 1,mean_sd)
mean_Ctrl20 <- apply(Fraction20_Ctrl, 1,mean_sd)
mean_Ctrl21 <- apply(Fraction21_Ctrl, 1,mean_sd)
mean_Ctrl22 <- apply(Fraction22_Ctrl, 1,mean_sd)
mean_Ctrl23 <- apply(Fraction23_Ctrl, 1,mean_sd)
mean_Ctrl24 <- apply(Fraction24_Ctrl, 1,mean_sd)
mean_Ctrl25 <- apply(Fraction25_Ctrl, 1,mean_sd)

mean_RNase1 <- apply(Fraction1_RNase, 1,mean_sd)
mean_RNase2 <- apply(Fraction2_RNase, 1,mean_sd)
mean_RNase3 <- apply(Fraction3_RNase, 1,mean_sd)
mean_RNase4 <- apply(Fraction4_RNase, 1,mean_sd)
mean_RNase5 <- apply(Fraction5_RNase, 1,mean_sd)
mean_RNase6 <- apply(Fraction6_RNase, 1,mean_sd)
mean_RNase7 <- apply(Fraction7_RNase, 1,mean_sd)
mean_RNase8 <- apply(Fraction8_RNase, 1,mean_sd)
mean_RNase9 <- apply(Fraction9_RNase, 1,mean_sd)
mean_RNase10 <- apply(Fraction10_RNase, 1,mean_sd)
mean_RNase11 <- apply(Fraction11_RNase, 1,mean_sd)
mean_RNase12 <- apply(Fraction12_RNase, 1,mean_sd)
mean_RNase13 <- apply(Fraction13_RNase, 1,mean_sd)
mean_RNase14 <- apply(Fraction14_RNase, 1,mean_sd)
mean_RNase15 <- apply(Fraction15_RNase, 1,mean_sd)
mean_RNase16 <- apply(Fraction16_RNase, 1,mean_sd)
mean_RNase17 <- apply(Fraction17_RNase, 1,mean_sd)
mean_RNase18 <- apply(Fraction18_RNase, 1,mean_sd)
mean_RNase19 <- apply(Fraction19_RNase, 1,mean_sd)
mean_RNase20 <- apply(Fraction20_RNase, 1,mean_sd)
mean_RNase21 <- apply(Fraction21_RNase, 1,mean_sd)
mean_RNase22 <- apply(Fraction22_RNase, 1,mean_sd)
mean_RNase23 <- apply(Fraction23_RNase, 1,mean_sd)
mean_RNase24 <- apply(Fraction24_RNase, 1,mean_sd)
mean_RNase25 <- apply(Fraction25_RNase, 1,mean_sd)

# Create seperate dataframe with mean for ctrl and RNase
mean_HeLa <- data.frame(mean_Ctrl1, mean_RNase1, mean_Ctrl2, mean_RNase2, mean_Ctrl3,mean_RNase3, mean_Ctrl4, mean_RNase4,  mean_Ctrl5, mean_RNase5, mean_Ctrl6, mean_RNase6, mean_Ctrl7, mean_RNase7, mean_Ctrl8, mean_RNase8, mean_Ctrl9, mean_RNase9, mean_Ctrl10, mean_RNase10, mean_Ctrl11, mean_RNase11, mean_Ctrl12, mean_RNase12, mean_Ctrl13, mean_RNase13, mean_Ctrl14, mean_RNase14, mean_Ctrl15, mean_RNase15, mean_Ctrl16, mean_RNase16, mean_Ctrl17, mean_RNase17, mean_Ctrl18, mean_RNase18, mean_Ctrl19, mean_RNase19, mean_Ctrl20, mean_RNase20, mean_Ctrl21, mean_RNase21, mean_Ctrl22, mean_RNase22, mean_Ctrl23, mean_RNase23, mean_Ctrl24, mean_RNase24, mean_Ctrl25, mean_RNase25)

HeLa_Ctrl_mean = mean_HeLa %>% select(contains("Ctrl"))
HeLa_RNase_mean = mean_HeLa %>% select(contains("RNA"))

# Remove excess objects
rm(mean_Ctrl1, mean_RNase1, mean_Ctrl2, mean_RNase2, mean_Ctrl3,mean_RNase3, mean_Ctrl4, mean_RNase4,  mean_Ctrl5, mean_RNase5, mean_Ctrl6, mean_RNase6, mean_Ctrl7, mean_RNase7, mean_Ctrl8, mean_RNase8, mean_Ctrl9, mean_RNase9, mean_Ctrl10, mean_RNase10, mean_Ctrl11, mean_RNase11, mean_Ctrl12, mean_RNase12, mean_Ctrl13, mean_RNase13, mean_Ctrl14, mean_RNase14, mean_Ctrl15, mean_RNase15, mean_Ctrl16, mean_RNase16, mean_Ctrl17, mean_RNase17, mean_Ctrl18, mean_RNase18, mean_Ctrl19, mean_RNase19, mean_Ctrl20, mean_RNase20, mean_Ctrl21, mean_RNase21, mean_Ctrl22, mean_RNase22, mean_Ctrl23, mean_RNase23, mean_Ctrl24, mean_RNase24, mean_Ctrl25, mean_RNase25)
```

## Calculate how many values failed one sigma cut-off

For the analysis, we quantified how many replicate values were excluded. 

```{r One sigma cut-off total}
# Function to test how many values fail one sigma cut-off
mean_sd_na <- function(x) {
        m <- mean(x)
        s <- sd(x)
       for (i in 1:3)
        if (( (m - 1 * s) > x[i]) || ((m + 1 * s) < x[i])) {x[i] <- 1}
        else (x[i] <- 0)
   sum(x)
}

# Select values outside of one sigma cut-off
na_mean_Ctrl1 <- apply(Fraction1_Ctrl, 1,mean_sd_na)
na_mean_Ctrl2 <- apply(Fraction2_Ctrl, 1,mean_sd_na )
na_mean_Ctrl3 <- apply(Fraction3_Ctrl, 1,mean_sd_na )
na_mean_Ctrl4 <- apply(Fraction4_Ctrl, 1,mean_sd_na )
na_mean_Ctrl5 <- apply(Fraction5_Ctrl, 1,mean_sd_na)
na_mean_Ctrl6 <- apply(Fraction6_Ctrl, 1,mean_sd_na)
na_mean_Ctrl7 <- apply(Fraction7_Ctrl, 1,mean_sd_na)
na_mean_Ctrl8 <- apply(Fraction8_Ctrl, 1,mean_sd_na)
na_mean_Ctrl9 <- apply(Fraction9_Ctrl, 1,mean_sd_na)
na_mean_Ctrl10 <- apply(Fraction10_Ctrl, 1, mean_sd_na)
na_mean_Ctrl11 <- apply(Fraction11_Ctrl, 1,mean_sd_na)
na_mean_Ctrl12 <- apply(Fraction12_Ctrl, 1,mean_sd_na)
na_mean_Ctrl13 <- apply(Fraction13_Ctrl, 1,mean_sd_na)
na_mean_Ctrl14 <- apply(Fraction14_Ctrl, 1,mean_sd_na)
na_mean_Ctrl15 <- apply(Fraction15_Ctrl, 1,mean_sd_na)
na_mean_Ctrl16 <- apply(Fraction16_Ctrl, 1,mean_sd_na)
na_mean_Ctrl17 <- apply(Fraction17_Ctrl, 1,mean_sd_na)
na_mean_Ctrl18 <- apply(Fraction18_Ctrl, 1,mean_sd_na)
na_mean_Ctrl19 <- apply(Fraction19_Ctrl, 1,mean_sd_na)
na_mean_Ctrl20 <- apply(Fraction20_Ctrl, 1,mean_sd_na)
na_mean_Ctrl21 <- apply(Fraction21_Ctrl, 1,mean_sd_na)
na_mean_Ctrl22 <- apply(Fraction22_Ctrl, 1,mean_sd_na)
na_mean_Ctrl23 <- apply(Fraction23_Ctrl, 1,mean_sd_na)
na_mean_Ctrl24 <- apply(Fraction24_Ctrl, 1,mean_sd_na)
na_mean_Ctrl25 <- apply(Fraction25_Ctrl, 1,mean_sd_na)

na_mean_RNase1 <- apply(Fraction1_RNase, 1,mean_sd_na)
na_mean_RNase2 <- apply(Fraction2_RNase, 1,mean_sd_na)
na_mean_RNase3 <- apply(Fraction3_RNase, 1,mean_sd_na)
na_mean_RNase4 <- apply(Fraction4_RNase, 1,mean_sd_na)
na_mean_RNase5 <- apply(Fraction5_RNase, 1,mean_sd_na)
na_mean_RNase6 <- apply(Fraction6_RNase, 1,mean_sd_na)
na_mean_RNase7 <- apply(Fraction7_RNase, 1,mean_sd_na)
na_mean_RNase8 <- apply(Fraction8_RNase, 1,mean_sd_na)
na_mean_RNase9 <- apply(Fraction9_RNase, 1,mean_sd_na)
na_mean_RNase10 <- apply(Fraction10_RNase, 1,mean_sd_na)
na_mean_RNase11 <- apply(Fraction11_RNase, 1,mean_sd_na)
na_mean_RNase12 <- apply(Fraction12_RNase, 1,mean_sd_na)
na_mean_RNase13 <- apply(Fraction13_RNase, 1,mean_sd_na)
na_mean_RNase14 <- apply(Fraction14_RNase, 1,mean_sd_na)
na_mean_RNase15 <- apply(Fraction15_RNase, 1,mean_sd_na)
na_mean_RNase16 <- apply(Fraction16_RNase, 1,mean_sd_na)
na_mean_RNase17 <- apply(Fraction17_RNase, 1,mean_sd_na)
na_mean_RNase18 <- apply(Fraction18_RNase, 1,mean_sd_na)
na_mean_RNase19 <- apply(Fraction19_RNase, 1,mean_sd_na)
na_mean_RNase20 <- apply(Fraction20_RNase, 1,mean_sd_na)
na_mean_RNase21 <- apply(Fraction21_RNase, 1,mean_sd_na)
na_mean_RNase22 <- apply(Fraction22_RNase, 1,mean_sd_na)
na_mean_RNase23 <- apply(Fraction23_RNase, 1,mean_sd_na)
na_mean_RNase24 <- apply(Fraction24_RNase, 1,mean_sd_na)
na_mean_RNase25 <- apply(Fraction25_RNase, 1,mean_sd_na)

# Data frame with values filtered out 
mean_HeLa_na <- data.frame (na_mean_Ctrl1, na_mean_RNase1, na_mean_Ctrl2, na_mean_RNase2, na_mean_Ctrl3,na_mean_RNase3, na_mean_Ctrl4, na_mean_RNase4,  na_mean_Ctrl5, na_mean_RNase5, na_mean_Ctrl6, na_mean_RNase6, na_mean_Ctrl7, na_mean_RNase7, na_mean_Ctrl8, na_mean_RNase8, na_mean_Ctrl9, na_mean_RNase9, na_mean_Ctrl10, na_mean_RNase10, na_mean_Ctrl11, na_mean_RNase11, na_mean_Ctrl12, na_mean_RNase12, na_mean_Ctrl13, na_mean_RNase13, na_mean_Ctrl14, na_mean_RNase14, na_mean_Ctrl15, na_mean_RNase15, na_mean_Ctrl16, na_mean_RNase16, na_mean_Ctrl17, na_mean_RNase17, na_mean_Ctrl18, na_mean_RNase18, na_mean_Ctrl19, na_mean_RNase19, na_mean_Ctrl20, na_mean_RNase20, na_mean_Ctrl21, na_mean_RNase21, na_mean_Ctrl22, na_mean_RNase22, na_mean_Ctrl23, na_mean_RNase23, na_mean_Ctrl24, na_mean_RNase24, na_mean_Ctrl25, na_mean_RNase25)

# Check how many values were removed
which(mean_HeLa_na == 3) # Three out of three values removed
which(mean_HeLa_na == 2) # Two out of three values removed
sum(mean_HeLa_na) # One out of three values removed

rm(na_mean_Ctrl1, na_mean_RNase1, na_mean_Ctrl2, na_mean_RNase2, na_mean_Ctrl3,na_mean_RNase3, na_mean_Ctrl4, na_mean_RNase4,  na_mean_Ctrl5, na_mean_RNase5, na_mean_Ctrl6, na_mean_RNase6, na_mean_Ctrl7, na_mean_RNase7, na_mean_Ctrl8, na_mean_RNase8, na_mean_Ctrl9, na_mean_RNase9, na_mean_Ctrl10, na_mean_RNase10, na_mean_Ctrl11, na_mean_RNase11, na_mean_Ctrl12, na_mean_RNase12, na_mean_Ctrl13, na_mean_RNase13, na_mean_Ctrl14, na_mean_RNase14, na_mean_Ctrl15, na_mean_RNase15, na_mean_Ctrl16, na_mean_RNase16, na_mean_Ctrl17, na_mean_RNase17, na_mean_Ctrl18, na_mean_RNase18, na_mean_Ctrl19, na_mean_RNase19, na_mean_Ctrl20, na_mean_RNase20, na_mean_Ctrl21, na_mean_RNase21, na_mean_Ctrl22, na_mean_RNase22, na_mean_Ctrl23, na_mean_RNase23, na_mean_Ctrl24, na_mean_RNase24, na_mean_Ctrl25, na_mean_RNase25)
```
It turns out, that only one out of three values was removed in 122175 triplicates.

Lastly, the names of the remaining proteins after noramlization were stored.
```{r Protein names}
proteinnames <- rownames(mean_HeLa)
```

## Check on normalization

Next, the results of the clean-up were compared to the raw data. We plotted the data pre- and post-normalization for six proteins. Three of which were sure "shifters" and three that were found not to be RNA-dependent.

"Shifters": SIN3A (transcriptional repressor), HDAC1 (Histone deacetylase 1), HNRPU (Heterogeneous nuclear ribonucleoprotein U)

"Non-shifters": ASNS (Asparagine synthetase), MCM2 (DNA replication licensing factor MCM2), MCM3 (DNA replication licensing factor MCM3)

```{r Comparing pre- and post-normalization}
fraction_index <- c(1:25)

# Repeat the following code block for the six test proteins ctrl vs RNase and pre vs post normalization

# Function to create comparison plots
compare_pre_post <- function(protein, y_limit_pre, y_limit_post) {

  par(mfrow = c(1,2))

  plot(fraction_index, HeLa_Ctrl1[protein,], type = 'l', col = 'forestgreen', # Control pre
      xlab = "Fractions",
      ylab = "Protein amount [arbitrary unit]",
      ylim = c(0,y_limit_pre),
      main = paste(protein, "pre-normalization"))
   lines(x = fraction_index, y = HeLa_RNase1[protein,], type = 'l', col = 'firebrick2') # RNase pre
   legend(x = "topright",                      # Position
       legend = c("ctrl", "RNase"),           # Legend texts
  col = c('darkgreen', 'red'),           # Line colors
       lwd = 2)                               # Line width

  plot(fraction_index, HeLa_Ctrl_mean[protein,], type = 'l', col = 'forestgreen', # Control post
     xlab = "Fractions",
     ylab = "Protein amount [%]",
     ylim = c(0,y_limit_post),
     main = paste(protein, "post-normalization"))
  lines(x = fraction_index, y = HeLa_RNase_mean[protein,], type = 'l', col = 'firebrick2') # RNase post
  legend(x = "topright",            
       legend = c("ctrl", "RNase"),           
       col = c('darkgreen', 'red'),         
       lwd = 2) 
}

# Apply function to our selected proteins

# Shifters
compare_pre_post("SIN3A_HUMAN", 4000000, 25) #SIN3A
compare_pre_post("HDAC1_HUMAN", 7500000, 25) #HDAC1
compare_pre_post("HDAC2_HUMAN", 7500000, 25) #HDAC2
compare_pre_post("RFC2_HUMAN", 3500000, 35) #RFC2

# Non-Shifters
compare_pre_post("ASNS_HUMAN", 2500000, 40) #ASNS
compare_pre_post("MCM2_HUMAN", 18000000, 25) #MCM2
compare_pre_post("MCM3_HUMAN", 25000000, 25) #MCM3

```

The scale represents percentages now with the area under the curve equaling 1. Furthermore, peaks should become more visible as well as "shifts".

# Shapiro-Wilk test

For future statistical tests, the normal distribution of our triplicate values was checked using the Shapiro-Wilk test.

```{r Shapiro-Wilk test}
# Define Shapiro test function
shapiro_testing <- function(x) {
  if (diff(range(x)) == 0) 
      list()
  else shapiro.test(x)
}

# Define function to apply shapiro testing to list of names for all fractions
shapiro_on_name_list <- function(list) {
  mean = c()
  for (i in 1:length(list)) {
    df = get(list[i])
    list_results <- apply(df, 1, shapiro_testing)
    p_values <- unlist(lapply(list_results, function(x) {x$p.value}))
    # print p mean value for each fraction
    mean = c(mean, mean(p_values))
  }
  mean
}

# Apply testing to data and test if all values are greater than 5 %
shapiro_values_greater_0.05 <- shapiro_on_name_list(fractions_names) > 0.05

# Remove excess variables
rm(Fraction1_Ctrl, Fraction1_RNase, Fraction2_Ctrl, Fraction2_RNase, Fraction3_Ctrl, Fraction3_RNase, Fraction4_Ctrl, Fraction4_RNase, Fraction5_Ctrl, Fraction5_RNase, Fraction6_Ctrl, Fraction6_RNase, Fraction7_Ctrl, Fraction7_RNase, Fraction8_Ctrl, Fraction8_RNase, Fraction9_Ctrl, Fraction9_RNase, Fraction10_Ctrl, Fraction10_RNase, Fraction11_Ctrl, Fraction11_RNase, Fraction12_Ctrl, Fraction12_RNase, Fraction13_Ctrl, Fraction13_RNase, Fraction14_Ctrl, Fraction14_RNase, Fraction15_Ctrl, Fraction15_RNase, Fraction16_Ctrl, Fraction16_RNase, Fraction17_Ctrl, Fraction17_RNase, Fraction18_Ctrl, Fraction18_RNase, Fraction19_Ctrl, Fraction19_RNase, Fraction20_Ctrl, Fraction20_RNase, Fraction21_Ctrl, Fraction21_RNase, Fraction22_Ctrl, Fraction22_RNase, Fraction23_Ctrl, Fraction23_RNase, Fraction24_Ctrl, Fraction24_RNase, Fraction25_Ctrl, Fraction25_RNase)
```

# Detection of global maxima

We identified the global maximum for wach protein both in the control and RNase-treated sample. 
```{r Find global maxima}
# Create function that selects all maxima by comparison of the protein content in a fraction with two neighboring values
find_peaks <- function (x, m = 2){
     shape <- diff(sign(diff(x, na.pad = FALSE)))
     pks <- sapply(which(shape < 0), FUN = function(i){
        z <- i - m + 1
        z <- ifelse(z > 0, z, 1)
        w <- i + m + 1
        w <- ifelse(w < length(x), w, length(x))
        if(all(x[c(z : i, (i + 2) : w)] <= x[i + 1])) return(i + 1) else return(numeric(0))
    })
     pks <- unlist(pks)
     pks
     
     if ( sum(x[1]>x[2:(m+1)]) == 2 ) {pks <- c(pks, 1)} 
     if ( sum(x[25]>x[24:(25-m)]) == 2 ) {pks <- c(pks, 25)}
     pks <- unlist(pks)
     pks
}

# Select global maxima for each protein and sample
absolute_max_ctrl <- apply(HeLa_Ctrl_mean, 1, max)
absolute_max_RNase <- apply(HeLa_RNase_mean, 1, max)

# Find fraction, where global maximum is located
fracs_ctrl <- max.col(HeLa_Ctrl_mean)
fracs_RNase <- max.col(HeLa_RNase_mean)

# Create data frame with global maximum value and fraction per protein and sample
abs_max_both <- data.frame(absolute_max_ctrl, fracs_ctrl, absolute_max_RNase, fracs_RNase)
```

# Calculate x-shift and direction

Next, we compared the x-positions of the global maxima in the control vs. RNase-treated sample.

```{r Calculate x-shift}
# Create for-loop that substracts fraction of global maximum of ctrl and RNase for each protein
i <- c(1:7081)
x_shifts <- c()
for(x in i){x_shifts[x] <- fracs_ctrl[x] - fracs_RNase[x]
}
# Add those values to existing data frame
abs_max_both$x_shifts <- x_shifts

# Create for-loop that calculates the difference of the global maximum fractions for ctrl and RNase for each protein
i <- c(1:7081)
y_shifts <- c()
for(x in i){y_shifts[x] <- absolute_max_ctrl[x] - absolute_max_RNase[x]
}

# Create for-loop that calculates the difference of the fraction of the global maximum ctrl and fraction values of RNase for each protein
i <- c(1:7081)
y_shifts2 <- c()
for(x in i){y_shifts2[x] <- absolute_max_ctrl[x] - HeLa_RNase_mean[x,fracs_ctrl[x]]
}

# Add those values to existing data frame
abs_max_both$y_shifts <- y_shifts
abs_max_both$y_shifts2 <- y_shifts2

# Define x-shift directions
no_shift <- which(abs_max_both$x_shifts == 0)
left_shift <- which(abs_max_both$x_shifts > 0)
right_shift <- which(abs_max_both$x_shifts < 0)

# Create for-loop to apply definitions for x-shift
i <- c(1:7081)
shift_direction <- c()
for(x in i){shift_direction[x] <- if(abs_max_both$x_shifts[x]==0) {'no shift'}
else if(abs_max_both$x_shifts[x] < 0) {'right shift'}
else {'left shift'}
}
# Add x-shift direction to existing data frame
abs_max_both$shift_direction <- shift_direction
```

# Statisitcal analysis

## F-test

Next, the varaince of our data was analyzed per protein per fraction ctrl vs. RNase. The calculated p-values were tested against the significance level of 0.01.

```{r Apply F-test}
# Different data frames for each sample
HeLa_Ctrl_Norm <- Norm_HeLa %>% select(contains("Ctrl"))
HeLa_RNase_Norm <- Norm_HeLa %>% select(contains("RNA"))

# Select every single protein (row)
F_test_p_values = c()
for (i in 1:nrow(HeLa_Ctrl_Norm)) {
  x <- as.numeric(HeLa_Ctrl_Norm[i,])
  y <- as.numeric(HeLa_RNase_Norm[i,])
  
  # Select triplicate values for each fraction
  for (k in 1:25) {  
    result <- var.test(x[c(k, (25+k), (50+k))], y[c(k, (25+k), (50+k))], alternative = "two.sided")
    new_p_value <- result$p.value
     # save p-value for every protein comparing ctrl vs RNase
     F_test_p_values <-  c(F_test_p_values, new_p_value)
  }
}

# Create p-values matrix
p_value_matrix <- matrix(data = F_test_p_values, byrow = TRUE, ncol = 25)

col_names_p <-  sapply(c(1:25), function(x) {
    str_glue(x)
})

# Create p-value data frame
p_value_df <- data.frame(p_value_matrix)
rownames(p_value_df) <-  row.names(HeLa_Ctrl_Norm)
colnames(p_value_df) <-  col_names_p

# Check ctrl maxima for p-value > 0.01
F_max_ctrl <- c()
for (i in 1:nrow(p_value_df)) {
  boolian <- p_value_df[i, abs_max_both$fracs_ctrl[i]] > 0.01
  F_max_ctrl <- c(F_max_ctrl, boolian)
}

# Check RNase maxima for p-value > 0.01
F_max_RNase <- c()
for (i in 1:nrow(p_value_df)) {
  boolian <- p_value_df[i, abs_max_both$fracs_RNase[i]] > 0.01
  F_max_RNase <- c(F_max_RNase, boolian)
}
list_F_max_RNase <- F_max_RNase
```

Proteins failing the F-test were filtered out, either because of FALSE or NaN/NA F-test results. Since those proteins could not be analyzed with the t-test, they will not play a role in determining the false-negative and true-negative rates.

```{r Quantification of F-test results}
# Selecting FALSE proteins
F_FALSE_ctrl <- abs_max_both[which(F_max_ctrl == FALSE), ]
F_FALSE_RNase <- abs_max_both[which(F_max_RNase == FALSE), ]

trash_F_FALSE <- unique(c(row.names(F_FALSE_ctrl), row.names(F_FALSE_RNase)))
length(trash_F_FALSE)

# Selecting NA proteins
F_NA_ctrl <- abs_max_both[which(is.na(F_max_ctrl)), ]
F_NA_RNase <- abs_max_both[which(is.na(F_max_RNase)), ]

trash_F_NA <- unique(c(row.names(F_NA_ctrl), row.names(F_NA_RNase)))
length(trash_F_NA)

# Removal of selected FALSE and Na proteins
abs_max_both_F <- abs_max_both[!rownames(abs_max_both) %in% trash_F_FALSE, ]
abs_max_both_F <- abs_max_both[!rownames(abs_max_both) %in% trash_F_NA, ]

# Number of proteins after F-Test
dim(abs_max_both_F)

# Create dataframe with F-Test positive proteins only
F_RNase_norm <- HeLa_RNase_Norm[!rownames(abs_max_both) %in% trash_F_FALSE, ]
F_RNase_norm <- HeLa_RNase_Norm[!rownames(abs_max_both) %in% trash_F_NA, ]

F_ctrl_norm <- HeLa_Ctrl_Norm[!rownames(abs_max_both) %in% trash_F_FALSE, ]
F_ctrl_norm <- HeLa_Ctrl_Norm[!rownames(abs_max_both) %in% trash_F_NA, ]
```
So 5881 proteins will be analyzed with the t-test.

## T-test

With the t-test, significant y-shifts were detected.

```{r t-test on data}
t_test_p_values <- c()

# Compare every row (protein) from Ctrl with corresponding RNase
for (i in 1:nrow(F_ctrl_norm)) {
  x <- as.numeric(F_ctrl_norm[i,])
  y <- as.numeric(F_RNase_norm[i,])
  
  for (k in 1:25) {  
    result <- t.test(x[c(k, (25+k), (50+k))], y[c(k, (25+k), (50+k))], alternative = "two.sided", var.equal = TRUE) # Comparison of triplicates
    new_p_value <- result$p.value
     # Save p-value for every protein comparing ctrl vs RNase
     t_test_p_values <-  c(t_test_p_values, new_p_value)
  }
}

# Matrix of p-values
p_value_matrix_t <- matrix(data = t_test_p_values, byrow = TRUE, ncol = 25)

col_names_p <-  sapply(c(1:25), function(x) {
    str_glue(x)
})

p_value_df_t <- data.frame(p_value_matrix_t)
rownames(p_value_df_t) <-  row.names(F_ctrl_norm)
colnames(p_value_df_t) <-  col_names_p

# For ctrl maxima
t_max_ctrl <- c()
for (i in 1:nrow(p_value_df_t)) {
  boolian <- p_value_df_t[i, abs_max_both_F$fracs_ctrl[i]] < 0.05 # Do t-values lay under the 0.05 threshold?
  t_max_ctrl <- c(t_max_ctrl, boolian)
}

list_t_max_ctrl <- t_max_ctrl

# For RNase maxima
t_max_RNase <- c()
for (i in 1:nrow(p_value_df_t)) {
  boolian <- p_value_df_t[i, abs_max_both_F$fracs_RNase[i]] < 0.05 # Do t-values lay under the 0.05 threshold?
  t_max_RNase <- c(t_max_RNase, boolian)
}

list_t_max_RNase <- t_max_RNase
```

Proteins failing the t-test were removed.

```{r Remove failed proteins}
# Remove NA
t_NA_ctrl <- abs_max_both_F[which(is.na(t_max_ctrl)), ]
t_NA_RNase <- abs_max_both[which(is.na(t_max_RNase)), ]

trash_t_NA <- c(row.names(t_NA_ctrl), row.names(t_NA_RNase))
length(trash_t_NA)

# Remove FALSE
t_FALSE_ctrl <- abs_max_both_F[which(t_max_ctrl == FALSE), ]
t_FALSE_RNase <- abs_max_both_F[which(t_max_RNase == FALSE), ]

trash_t_FALSE <- unique(c(row.names(t_FALSE_ctrl), row.names(t_FALSE_RNase)))
length(trash_t_FALSE)

abs_max_both_t <- abs_max_both[!rownames(abs_max_both_F) %in% trash_t_FALSE, ]
dim(abs_max_both_t) #dataset without FALSE

t_RNase_norm <- F_RNase_norm[!rownames(abs_max_both_F) %in% trash_t_FALSE, ]
t_ctrl_norm <- F_ctrl_norm[!rownames(abs_max_both_F) %in% trash_t_FALSE, ]

#Remove proteins with all zero triplicates
trash_t_y0 <- c(rownames(abs_max_both_t[which(abs_max_both_t$y_shifts == 0), ]))

#Cleaned data frame
abs_max_both_t_y0 <- abs_max_both_t[!rownames(abs_max_both_t) %in% trash_t_y0, ]
nrow(abs_max_both_t_y0)
```

## Methods for Identification of potential RBPs

For the next analysis, we further filtered out the significant y-shifters that showed no significant x-shift (shift > 1).
```{r x shift > 1}
# At least 1 or more fractions x shift
trash_no_x_1 <- c(rownames(abs_max_both_t_y0[which(abs(abs_max_both_t_y0$x_shifts) < 2), ]))
abs_max_both_x_y_1 <- abs_max_both_t_y0[!rownames(abs_max_both_t_y0) %in% trash_no_x_1, ]
abs_max_both_x_y_1

# All proteins that entered analysis
All_potential_candidates <- row.names(F_ctrl_norm)

# Results t-test
RBP_t <- c(rownames(abs_max_both_x_y_1)) # Proteins passing t-test = RBP
Non_RBP_t <- All_potential_candidates[!All_potential_candidates %in% RBP_t] # Proteins failing t-test = non-RBP
length(RBP_t)
length(Non_RBP_t)
```

We created functions to evaluate our results. They show how many of the detected RBPs or non-RBPs were identified previously in other publications. Furthermore, we wanted to check, if false-positive RBPs were at least RNA binding dependent (RBD).
```{r Quantification of results functions}
# Function to identify True RBP
Find_true_RBP <- function(potential_RBP) {
  Match_RBP <- HS_RBPs$Entry_Name.x %in% potential_RBP    # Compare potential RBP with RBP from other publications
  True_RBP_pos <- which(Match_RBP == TRUE)                # Find the positions of true RBP
  True_RBP_names <- HS_RBPs$Entry_Name.x[True_RBP_pos]    # Find names of true RBP
  return(True_RBP_names)
}

# Function to identify False RBP
Find_false_RBP <- function(potential_RBP) {                    
  Match_RBP <- HS_non_RBPs$Entry_Name.x %in% potential_RBP      # Compare potential RBP with RBP from other publications
  False_RBP_pos <- which(Match_RBP == TRUE)                     # Find the positions of false RBP
  False_RBP_names <- HS_non_RBPs$Entry_Name.x[False_RBP_pos]    # Find names of false RBP
  return(False_RBP_names)
}

# Function to identify true non-RBP
Find_true_non_RBP <- function(potential_non_RBP) {
  Match_non_RBP <- HS_non_RBPs$Entry_Name.x %in% potential_non_RBP    # Compare potential non-RBP with non-RBP from other publications
  True_non_RBP_pos <- which(Match_non_RBP == TRUE)                    # Find the positions of true non-RBP
  True_non_RBP_names <- HS_non_RBPs$Entry_Name.x[True_non_RBP_pos]    # Find names of true non-RBP
  return(True_non_RBP_names)
}

# Function to identify false non-RBP
Find_false_non_RBP <- function(potential_non_RBP) {
  Match_non_RBP <- HS_RBPs$Entry_Name.x %in% potential_non_RBP    # Compare potential non-RBP with non-RBP from other publications
  False_non_RBP_pos <- which(Match_non_RBP == TRUE)               # Find the positions of false non-RBP
  False_non_RBP_names <- HS_RBPs$Entry_Name.x[False_non_RBP_pos]   # Find names of false non-RBP
  return(False_non_RBP_names)
}
# Function to show RBDs
Find_RBD <- function(potential_RBD) {
  RBD_stat <- RBD_stat_non_RBP[rownames(RBD_stat_non_RBP) %in% potential_RBD, ]
  RBD_true_prot_df <- RBD_stat[RBD_stat$RBD_status == "RBD", ]
  RBD_true_prot_names <- rownames(RBD_true_prot_df)
  return(RBD_true_prot_names)
}
```

## Sum up t-test results

The new functions were applied to evaluate the t-test.
```{r Quantification t-test results}
# True RBP
True_RBP_t <- Find_true_RBP(RBP_t)
length(True_RBP_t)

# False RBP
False_RBP_t <- Find_false_RBP(RBP_t)
length(False_RBP_t)

# True non-RBP
True_non_RBP_t <- Find_true_non_RBP(Non_RBP_t)
length(True_non_RBP_t)

# False non-RBP
False_non_RBP_t <- Find_false_non_RBP(Non_RBP_t)
length(False_non_RBP_t)

# Show RBD status of all non-RBP
RBD_stat_non_RBP <- HS_non_RBPs[ ,c("Entry_Name.x", "RBD_status")]
row.names(RBD_stat_non_RBP) <- HS_non_RBPs$Entry_Name.x

# Specificity and Sensitivity parameters
( length(True_RBP_t) / nrow(abs_max_both_x_y_1) ) * 100 # True positive quote
( length(False_RBP_t) / nrow(abs_max_both_x_y_1) ) * 100 # False positive quote

( length(True_non_RBP_t) / length(Non_RBP_t) ) * 100 # True negative quote
( length(False_non_RBP_t) / length(Non_RBP_t) ) * 100 # False negative quote

# Remove excess objects
rm(abs_max_both_F, abs_max_both_t, F_ctrl_norm, F_FALSE_ctrl, F_FALSE_RNase, F_NA_ctrl, F_NA_RNase, F_RNase_norm,p_value_df_t, p_value_matrix_t, t_ctrl_norm, t_FALSE_ctrl, t_FALSE_RNase, t_NA_ctrl, t_NA_RNase, t_RNase_norm, p_value_matrix, result)

rm(HeLa_cleaned, HeLa_cleaned_NA)
```


# Clustering

First, we derived clustering variables from our data.
```{r Clustering variables}
# Define function to calculate quotient for clustering variables
quoti <- function(data1, data2){
  
Quotienten = c()
  for (k in 1:7081) {
    dat = data1[k,]
    dat2 = data2[k,]
    Quotient = c()
    
      for (i in 1:25) {
        if (dat2[i] == 0)  { quotient <- 0 }
        else quotient <- dat[i]/dat2[i]
        Quotient = c(Quotient, quotient)
      }
      Quotienten = rbind(Quotienten, Quotient)
  } 
  rownames(Quotienten) = rownames( data1)
  return(Quotienten)
}

# Select highest quotient sum values
# We checked the quotient sum value of each protein and selected only those above 20 (random value). Those above 20, seemed to contain many RBP's compared to the RDeep database. 

quotienten_HeLa_RNase <- quoti(HeLa_RNase_mean, HeLa_Ctrl_mean)
q_HeLa_mat <- matrix(unlist(quotienten_HeLa_RNase), ncol=25, byrow=FALSE)

# Variables derived from quotient
maxi_RNase <- apply(q_HeLa_mat,1, max)
mean <- rowMeans(q_HeLa_mat)
x_abs <- abs(x_shifts)
q_mat_sums <- rowSums(q_HeLa_mat, na.rm=TRUE)
abs_mat_both_q <- cbind(abs_max_both, maxi_RNase, x_abs, mean,q_mat_sums)
```

With the new variables, we created three clusters to predict RBPs.

```{r Clustering}
# Create cluster data frame
cluster_data <- data.frame(maxi_RNase, q_mat_sums)

# Calculate optimal number of clusters
fviz_nbclust(cluster_data, kmeans, method='silhouette')

# Cluster
kmeans_fancy <- kmeans(scale(cluster_data), 3, nstart = 25)
kmeans_fancy

# Plot cluster
fviz_cluster(kmeans_fancy, data = scale(cluster_data), geom = c("point"),ellipse.type = "euclid")

# Silhouette plot for cluster quality
sil <- silhouette(kmeans_fancy$cluster, dist(cluster_data))
fviz_silhouette(sil)
```

## Clustering results

First, we created vectors containing the assigned protein names of each cluster.
```{r Resulting cluster assingment}
# Extract information on cluster centers
cluster_centers <- data.frame(kmeans_fancy$centers)
cluster_centers

# Cluster A closest to origin, cluster C farthest away
Cluster_C_number <- which.max(cluster_centers$maxi_RNase)
Cluster_A_number <- which.min(cluster_centers$maxi_RNase)
Cluster_B_number <- (1+2+3) - Cluster_C_number - Cluster_A_number

# Cluster content
Cluster_C <- proteinnames[which(kmeans_fancy$cluster==Cluster_C_number)]
Cluster_B <- proteinnames[which(kmeans_fancy$cluster==Cluster_B_number)]
Cluster_A <- proteinnames[which(kmeans_fancy$cluster==Cluster_A_number)]

# All tested
Cluster_total <- c(Cluster_C, Cluster_B, Cluster_A)
```

### Cluster C
Cluster C was supposed to be a safe "true RBP" cluster. This was tested:
```{r Cluster C Quantification}
# True RBP
True_RBP_C <- Find_true_RBP(Cluster_C)
length(True_RBP_C)

# False RBP
False_RBP_C <- Find_false_RBP(Cluster_C)
length(False_RBP_C)

#RBD status
False_RBP_true_RBD_C <- Find_RBD(False_RBP_C)
length(False_RBP_true_RBD_C)

# Senstivity and Specificity markers
( length(True_RBP_C) / length(Cluster_C) ) * 100 # True positive quote
( length(False_RBP_C) / length(Cluster_C) ) * 100 # False positive quote
```
Cluster C included mostly true RBPs.

### Cluster B  

We anticipated Cluster B to suggest more RBP, although with reduced precision.
```{r Cluster B Quantification}
# True RBP
True_RBP_B <- Find_true_RBP(Cluster_B)
length(True_RBP_B)

# False RBP
False_RBP_B <- Find_false_RBP(Cluster_B)
length(False_RBP_B)

#RBD status
False_RBP_true_RBD_B <- Find_RBD(False_RBP_B)
length(False_RBP_true_RBD_B)

#Senstivity and Specificity markers
( length(True_RBP_B) / length(Cluster_B) ) * 100 # True positive quote
( length(False_RBP_B) / length(Cluster_B) ) * 100 # False positive quote
```
Cluster B included mostly true RBPs.

### Cluster A

We anticipated Cluster A to suggest mostly non-RBP.
```{r Cluster A Quantification}
# True non-RBP
True_non_RBP_A <- Find_true_non_RBP(Cluster_A)
length(True_non_RBP_A)

# False non-RBP
False_non_RBP_A <- Find_false_non_RBP(Cluster_A)
length(False_non_RBP_A)

#Senstivity and Specificity markers
( length(True_non_RBP_A) / length(Cluster_A) ) * 100 # True positive quote
( length(False_non_RBP_A) / length(Cluster_A) ) * 100 # False positive quote
```

Cluster A included both several non-RBP and RBP.


### Summarizing Cluster B + C

Cluster B and C were combined because they both predicted RBPs accurately. 
```{r Cluster BC Quantification}
Cluster_BC <- c(Cluster_B, Cluster_C) # Combined cluster BC
length(Cluster_BC)

# True RBP
True_RBP_BC <- Find_true_RBP(Cluster_BC)
length(True_RBP_BC)

# False RBP
False_RBP_BC <- Find_false_RBP(Cluster_BC)
length(False_RBP_BC)

#RBD status
False_RBP_true_RBD_BC <- Find_RBD(False_RBP_BC)
length(False_RBP_true_RBD_BC)

# Senstivity and Specificity markers
( length(True_RBP_BC) / length(Cluster_BC) ) * 100 # True positive quote
( length(False_RBP_BC) / length(Cluster_BC) ) * 100 # False positive quote
```

## Plotting Shifters

Based on the x-shifts, we plotted our proteins as three groups: Those with x-shift = 0, x-shift > 0 or x-shift < 0. 

```{r Plot shifters}
# Assign numbers for shift behavior
i <- c(1:7081)
ID <- c()
for(x in i){
ID[x] <- if (abs_mat_both_q$x_shifts[x] == 0){abs_mat_both_q$ID[x] =1 # x-shift = 0
} else if(abs_mat_both_q$x_shifts[x] > 0){abs_mat_both_q$ID[x] =2 # x-shift > 0
} else {abs_mat_both_q$ID[x] =3} # x-shift < 0 
}
# Plot all proteins
control_fraction <- as.matrix(abs_mat_both_q$fracs_ctrl)
RNase_fraction <- as.matrix(abs_mat_both_q$fracs_RNase)
cols <- c(1,3)

colors <- c("right_shifters" = "green", "left_shifters" = "red", "non_shifters" = "black")

p_all <- ggplot(abs_mat_both_q[,cols], aes(x = RNase_fraction, y=control_fraction, color = ID, show.legend = TRUE)) +
geom_point(color = ID) +
  geom_density_2d(alpha = 0.5)+
  scale_colour_manual(values = colors)
   
p0 <- p_all + labs(title = "All proteins")
p0
```

#Linear regression 

Lastly, we used our data for a linear regression model.

## Train the model

We used 5000 randomly chosen proteins to train the model.
```{r Linear regression model}
# Select 5000 proteins from raw data randomly
proteins <- c(1:7081)
samples.numbers.training <- sample(proteins,5000)
samples.numbers.test <- proteins[-c(samples.numbers.training)]

random.samples.training <- abs_mat_both_q[samples.numbers.training,c(5,9,12)]
colnames(random.samples.training) <- c("x","max","sum")

random.samples.test <- abs_mat_both_q[samples.numbers.test,c(5,9,12)]
colnames(random.samples.test) <- c("x","max","sum")

modell <- lm(x ~ sum + max  , data = random.samples.training )
summary(modell)

plot(random.samples.training$x,random.samples.training$max, col ="blue", abline(lm(x ~ max, data = random.samples.training )), ylim = c(0,15), xlab = "x-shift", ylab = " Maximum of quotients", main = "Regression value Maxima")
plot(random.samples.training$x,random.samples.training$sum, col ="blue", abline(lm(x ~ sum, data = random.samples.training )), ylim = c(0,15), xlab = "x-shift", ylab = "Sums of quotients", main = "Regression value Sum")

``` 

## Test the model

We then tested the model on the remaining 2081 proteins.
```{r Testing regression model}
# Use remaining 2081 proteins to test the regression model
c <- predict(modell, newdata = random.samples.test)

i <- c(1:2081)
shift_direction.regression <- c()
for(x in i){shift_direction.regression[x] <- if(c[x] < 1) {'no shift'
}else  {'shift'}
}

shifters <- x_shifts[samples.numbers.test]
shift_direction.code <- c()
for(x in i){shift_direction.code[x] <- if(shifters[x] == 0) {'no shift'
}else  {'shift'}
}

compare <- cbind(c, x_abs[samples.numbers.test], shift_direction.regression, shifters, shift_direction[samples.numbers.test])

```


## Results Regression

Now we will compare the results of the linear regression with the additional data.
```{r Quantification of regression results}
Regression_data <- data.frame(compare)

RBP_regression <- proteinnames[which(Regression_data$shift_direction.regression == "shift")] # identified RBP
non_RBP_regression <- proteinnames[which(Regression_data$shift_direction.regression == "no shift")] # indentified non-RBP

# Number of RBP and non-RBP identified
length(RBP_regression)
length(non_RBP_regression)

# True RBP
True_RBP_reg <- Find_true_RBP(RBP_regression)
length(True_RBP_reg)

# False RBP
False_RBP_reg <- Find_false_RBP(RBP_regression)
length(False_RBP_reg)

# True non-RBP
True_non_RBP_reg <- Find_true_non_RBP(non_RBP_regression)
length(True_non_RBP_reg)

# False non-RBP
False_non_RBP_reg <- Find_false_non_RBP(non_RBP_regression)
length(False_non_RBP_reg)

#RBD-status
False_RBP_true_RBD_reg <- Find_RBD(False_RBP_reg)
length(False_RBP_true_RBD_reg)

# Quantification
( length(True_RBP_reg) / length(RBP_regression) ) * 100 # True positive quote
( length(False_RBP_reg) / length(RBP_regression) ) * 100 # False positive quote

( length(True_non_RBP_reg) / length(non_RBP_regression) ) * 100 # True negative quote
( length(False_non_RBP_reg) / length(non_RBP_regression) ) * 100 # False negative quote
```

## Check on 7 previously selected control proteins

Lastly, we wanted to check, if the methods were able to correctly identify our selected shifters and non-shifters.

```{r Checking on seven selected proteins}
# Shifters
shifters <- c("SIN3A_HUMAN", "HDAC1_HUMAN", "HDAC2_HUMAN", "RFC2_HUMAN")
# Non-Shifters
non_shifters <- c("ASNS_HUMAN", "ASNS_HUMAN", "MCM3_HUMAN")

# t-test
table(shifters %in% RBP_t)
table(non_shifters %in% Non_RBP_t)
table(non_shifters %in% RBP_t) # non-shifters in shifters?
non_shifters[non_shifters %in% RBP_t]

# Clustering
table(shifters %in% Cluster_BC)
table(non_shifters %in% Cluster_A)
table(non_shifters %in% Cluster_BC) # non-shifters in shifters?

# Regression
table(shifters %in% RBP_regression)
table(non_shifters %in% non_RBP_regression)

```

## Analyze overlap of all methods

We compared whether all methods identified the same RBPs and non-RBPs or different ones.
```{r Overlap three methods}
# Results identification via statistical tests
i <- c(1:7081)
shift_direction.identification <- c()
for(x in i){shift_direction.identification[x] <- if(rownames(abs_mat_both_q)[x] %in% Non_RBP_t) {'no shift'
}else  {'shift'}
}

# Results clustering
i <- 1:7081
shift_direction.clustering <- c()
for(x in i){shift_direction.clustering[x] <- if(rownames(abs_mat_both_q)[x] %in% Cluster_A) {'no shift'
}else  {'shift'}
}

# Results regression
all.samples <- abs_mat_both_q[,c(5,9,12)]
colnames(all.samples) <- c("x","max","sum")

results_regression <- predict(modell, newdata = all.samples)

i <- c(1:7081)
shift_direction.regression.all <- c()
for(x in i){shift_direction.regression.all[x] <- if(results_regression[x] < 1) {'no shift'
}else  {'shift'}
}

overlap <- data.frame(shift_direction.identification,shift_direction.clustering, shift_direction.regression.all)

# Compare two methods with one another
# Statistical test vs. Clustering
compare.identification.clustering <-(shift_direction.identification[i] == shift_direction.clustering[i]) 
length (which(compare.identification.clustering == TRUE))

# Statistical test vs. Regression
compare.identification.regression <- (shift_direction.identification[i] == shift_direction.regression.all[i])
length (which(compare.identification.regression == TRUE))

# Regression vs. Clustering
compare.regression.clustering <- (shift_direction.regression.all[i] == shift_direction.clustering[i])
length (which(compare.regression.clustering == TRUE))

# Compare all three methods with one another
compare.shifts.all <- c() 
for (x in 1:7081) {
 a <- (shift_direction.identification[x] == shift_direction.clustering[x] & shift_direction.clustering[x] == shift_direction.regression.all[x])
 compare.shifts.all <- c(compare.shifts.all, a)
}

overlap <- cbind(overlap, compare.identification.clustering, compare.identification.regression, compare.regression.clustering, compare.shifts.all)

length (which(compare.shifts.all == TRUE))
```
The same 5259 Proteins were identified as RBPs or non-RBPs respectively using statistical tests and clustering.
The same 3310 Proteins were identified as RBPs or non-RBPs respectively using statistical tests and regression.
The same 4470 Proteins were identified as RBPs or non-RBPs respectively using regression and clustering.

All three methods identified the same 2979 RBPs or non-RBPs respectively.

## Cross-reference overlap of all three methods
```{r All tests overlap}
#select proteinnames of overlap
over.prot <- which(compare.shifts.all == TRUE)
overlap_proteins <- proteinnames[over.prot]

overlap.potential.non.RBP <-overlap_proteins[results_regression < 1]
overlap.potential.RBPn <- overlap_proteins[results_regression > 1]

# True RBP
True_RBP_overlap <- Find_true_RBP(overlap.potential.RBPn)
length(True_RBP_overlap)

# False RBP
False_RBP_overlap <- Find_false_RBP(overlap.potential.RBPn)
length(False_RBP_overlap)

# True non-RBP
True_non_RBP_overlap <- Find_true_non_RBP(overlap.potential.non.RBP)
length(True_non_RBP_overlap)

# False non-RBP
False_non_RBP_overlap <- Find_false_non_RBP(overlap.potential.non.RBP)
length(False_non_RBP_overlap)
``` 
The overlap of all three methods consists of 728 true RBP, 483 false RBP, 715 true non-RBP and 1032 false RBP.

## Clustering vs Statistical Tests

We compared the results of the clustering and the statistical analysis, because we discovered the greatest overlap between any pairs of applied methods.
```{r Overlap Clustering vs Statistical tests}
#select overlap proteinnames 
over.prot.id.clus <- which( compare.identification.clustering == TRUE)
overlap_proteins.id.clus <- proteinnames[over.prot.id.clus]

i <- 1:5240
overlap_potential.id.clus <- c()
for(x in i){overlap_potential.id.clus[x] <- if(overlap_proteins.id.clus[x] %in% Cluster_A) {TRUE
}else  {FALSE}
}

overlap_potential.non.RBP.id.clus <- overlap_proteins.id.clus[c(overlap_potential.id.clus == TRUE)]
overlap_potential.RBP.id.clus <- overlap_proteins.id.clus[c(overlap_potential.id.clus == FALSE)]

# True RBP
True_RBP_overlap <- Find_true_RBP(overlap_potential.RBP.id.clus)
length(True_RBP_overlap)

# False RBP
False_RBP_overlap <- Find_false_RBP(overlap_potential.RBP.id.clus)
length(False_RBP_overlap)

# True non-RBP
True_non_RBP_overlap <- Find_true_non_RBP(overlap_potential.non.RBP.id.clus)
length(True_non_RBP_overlap)

# False non-RBP
False_non_RBP_overlap <- Find_false_non_RBP(overlap_potential.non.RBP.id.clus)
length(False_non_RBP_overlap)
```
The overlap of clustering and statistical test results consists of 182 true RBP, 9 false RBP, 1623 true non-RBP and 3415 false RBP.